### 2016-05-28 Paper/paper.tex
more paper edits

\section{Introduction}
\citet{Sykulski2016} present\vmdel{s} a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
Figure~\ref{fig: fig1} shows a sample 250-day drifter trajectory. 
Oceanographers use these time series to increase understanding of  ocean circulation patterns.
Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper makes a notable contribution to the statistical literature by illustrating how to directly embed knowledge of a physical system into a stochastic model. 
 \begin{figure}[h!]
  \centering
    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue. The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
        	\label{fig: fig1}
\end{figure}

\par
The motion of parcels of water moving over changing latitudes is known to have a rotational component. 
Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.  
This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude. 
Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions of rotation in the two hemispheres.
 Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
  However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
 The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citet{Isern2003} or \citet{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
\citet{Shoosmith2005} and \citet{Lankhorst2006} do identify eddies by finding periods of repeated looping in the drifters' trajectories with a simple autoregressive model.
 Similarly, \citet{Boebel2003} calculate the curvature of the trajectory  and then classify eddies as periods of consistent rotations in one direction. 
 \citet{Sykulski2016} assert that they are the first to identify eddies by explicitly modeling inertial oscillations within a stochastic model.  
 \par
In achieving this result, several statistical advances are made.  
The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations. 
The Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is applied for the first time to model the turbulent background. 
Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations. 
Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
\par
The remainder of this report proceeds as follows.
 After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it. 
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model. 
We conclude the paper with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.

@@

	\subsection{Spectral Analysis}
	\label{sec: specAnalysis}
	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.  
	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately weighted complex exponentials. 
	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations:
	\hdcomment{these are wrong} 	
	\begin{align}
z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega,&& f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
\end{align}
\citep{Percival1993}. 
To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
\begin{equation}
S_{z}(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
\end{equation}
is often employed. 
This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
We define the autocovariance  for $z(t)$ as $s_{z}(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
\begin{align}
\label{eq: fourierPair}
S_{z}(\omega) = \int_{-\infty}^{\infty}s_{z}(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s_{z}(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S_{z}(\omega) e^{i \omega t} d\omega 
\end{align}

\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, where 
\begin{equation}
\label{eq: perio}
\hat{S}_{z}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t} \right)^{2}
\end{equation}
\citep{Sykulski2013}. 
Although intuitively appealing, this estimator has known problems that stem from  sampling a time series at discrete intervals. 
\textbf{\it{Aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.  
\hdcomment{Get this right!} Additionally, \textbf{\it{Leakage}} refers to the effects of because Fourier transforms are applied are a smooth, create discontinuities
$\textbf{\it{Tapering}}$ reduces leaking by \hdcomment{get this right!}
However, depending on how tapering is used, it can introduce its own  biases in estimation.

\subsection{Stochastic Model}
 To model a drifter's movement over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, is converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$. 
 Separate stochastic models are introduced for the inertial oscillation and turbulent background components of the drifter's behavior. 
 Working in the frequency domain, the two components of the model are added together, creating a six-parameter spectral model. 
 Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood, a relatively new approximation to the true Whittle likelihood that compensates for bias and handles aliasing \citep{Sykulski2013}.

\subsubsection{Inertial oscillations}
We begin by discussing the model for the inertial oscillations. 
In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations
\begin{align}
\label{eq: deterOsc}
\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu,\\ \nonumber
\frac{\partial v}{\partial t} - f_{0}u &= G - cv,
\end{align}
where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}. 
Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows
\begin{align}
\label{eq:diffEqDeriv}
\nonumber
\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt} \\ \nonumber
dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
&= f_{0}(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
&= if_{0}(u + iv)dt - c(u + iv)dt + (F + iG)dt\\ 
&= (if_{0} - c)z dt + (F + iG)dt. 
\end{align}
\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
 \begin{equation}
\label{eq: ouEq}
dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
\end{equation}  
Equation ~\ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
In general,  OU processes are considered continuous-time equivalents to the AR(1) process. Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate Markov process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
 The complex-valued OU process is the complex representation of a bivariate OU process.  
 In the formulation of Equation~\ref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
 The final term of Equation~\ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.  The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
 We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $\tau$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation~\ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data. 
If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.

For time series modeling, we are interested in the autocovariance of this process at different time lags. To derive the autocovariance as given as Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ is defined as 
\begin{equation}
\label{eq: expec}
s(\tau) = \mathbb{E}[z(t) z^{*}(t + \tau)] = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i(\mathbb{E}[u(t) v(t + \tau)] -\mathbb{E}[u(t + \tau)v( t)])
\end{equation}
with time lag $\tau >0$ \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
\begin{align}
\left( \begin{array}{c} du(t) \\ dv(t) \end{array} \right) &= 
\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q_{1}(t) \\ Ad Q_{2}(t) \end{array} \right). 
\end{align}
For a bivariate OU process, the covariance matrix at time lag $\tau > 0$ is 
\begin{align}
\label{eq: bivOU}
C(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}C(0)
\end{align}
\citep{Schach1971}.  \citet{Sykulski2016} specify that at the initial time, $z(t) \sim$ N(0, $A^{2}/2c$). We also note $u(t)$ and $v(t)$ are uncorrelated at the initial time, giving $\mathbb{E}[u(t)v(t)] = \mathbb{E}[v(t) u(t)] = 0$. Thus, Equation \ref{eq: expec} simplifies to
\begin{equation}
\label{eq: s0}
s(0) = \frac{A^{2}}{2c} =  \mathbb{E}[u(t)^{2}] + \mathbb{E}[(v(t)^{2}]. 
\end{equation}
Since $u(t)$ and $v(t)$ are assumed to be identical processes, Equation \ref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation ~\ref{eq: bivOU} can be expressed as 
\begin{align}
\label{eq: covDeriv}
 C(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &= 
\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right) 
\end{align}
where the second equality can be seen by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
Finally, using the values in Equation \ref{eq: covDeriv} in Equation \ref{eq: expec}, we obtain the autocovariance  as defined in \citet{Sykulski2016}
\begin{align}
\label{eq: ouAC} \nonumber
s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0}) \\
&= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
\end{align}
 The $(o)$ superscript denotes that this equation is for the OU component of the full model. Using Equation ~\ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
\label{eq:ouPSD}
S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
\end{align}

@@

Towards this end, the authors propose using a Mat\'{e}rn covariance structure \citep{Gneiting2012} for the turbulent background. 
The Mat\'{e}rn model has gained prominence in the spatial statistics literature due to its flexibility.
 In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. 
 The Mat\'{e}rn model is defined either by its autocovariance 
\begin{align}
\label{eq:maternAC}
s^{(m)}(\tau) = \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\mathcal{K}_{\alpha - 1/2}(h|\tau|),
\end{align}
or power spectral density,
\begin{align}
\label{eq:maternPSD}
S^{(m)}(\omega) = \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}},
\end{align}
where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\mathcal{K}_{\eta}$ is a modified Bessel function of the second kind of order $\eta$  and the superscript $(m)$ indicates that this is the Mat\'{e}rn part of the model \citep{Stein2012}. 
The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
 The parameter $\alpha$ is of particular importance, since it controls the smoothness, or degree of differentiability, of the process \citep{Fuentes2010}.

The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model. 
This means that these models can still be obtained under the current model if model fitting leads to particular parameter estimate values. 
They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
 Figure~\ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{ReplicatedFigures/fig2.pdf}
        \caption{Periodograms of the highlighted green and blue sections of the trajectories in Figure ~\ref{fig: fig1}. The inertial frequencies are marked in red. }
        	\label{fig: fig2}
\end{figure}

>>>>

\section{Introduction}
\citet{Sykulski2016} present a multi-component spectral model for analysis of data transmitted by ocean surface drifters. 
Drifters are free-floating  instruments that transmit their location at regular time intervals,  creating a \textbf{\it{Langrangian time series}}, or a sequence of spatial locations over time. 
Figure \ref{fig: fig1} shows a sample 250-day drifter trajectory. 
Oceanographers use these time series to increase understanding of  ocean circulation patterns.
Published in January 2016 in the \textbf{\it{Journal of the Royal Statistical Society Series C}}, this paper makes a notable contribution to the statistical literature by illustrating how to directly embed knowledge of a physical system into a stochastic model. 
 \begin{figure}[h!]
  \centering
    \includegraphics[width=.6\textwidth]{ReplicatedFigures/fig1.pdf}
        \caption{A 250-day sample drifter trajectory with two 50-day periods highlighted in green and blue. The time period highlighted in green is not affected by an eddy while the time period highlighted  in blue is.}
        	\label{fig: fig1}
\end{figure}

\par
The motion of parcels of water moving over changing latitudes is known to have a rotational component. 
Because the Earth has  different diameter at different latitudes, the speed of objects moving with Earth's rotation varies across latitudes.  
This induces a phenomena known as the Coriolis Effect, whereby very fast moving objects or objects observed over long periods, such as drifters, end up moving at speeds that are different than the Earth beneath them as they change latitude. 
Viewed from a stationary reference frame, this creates circular motions, or inertial oscillations. 
These motions have known frequencies, referred to as inertial frequencies, which take on positive values in the Southern hemisphere and negative values in the Northern hemisphere corresponding to the different directions objects rotate in the two hemispheres.
 Oceanographers  are interested in detecting deviations from these frequencies, as they are thought to indicate eddies, or persistent circular wave patterns \citep{Kunze1985}.
  However, the ocean has a constant turbulent background flow which makes identifying eddies in real data difficult \citep{ Elipot2010}.
 The majority of the previous work in oceanography has focused on detecting eddies using satellite altimetry rather than drifter data (see \citet{Isern2003} or \citet{Fu2010}, for example), so there is limited work on detecting eddies from drifters.
\citet{Shoosmith2005} and \citet{Lankhorst2006} do identify eddies by finding periods of repeated looping in the drifters' trajectories with a simple autoregressive model.
 Similarly, \citet{Boebel2003} calculate the curvature of the trajectory  and then classify eddies as periods of consistent rotations in one direction. 
 \citet{Sykulski2016} assert that they are the first to identify eddies by explicitly modeling inertial oscillations within a stochastic model.  
 \par
In achieving this result, several statistical advances are made.  
The authors introduce an additive spectral domain model that has components corresponding to both the turbulent background and the inertial oscillations. 
The Mat\'{e}rn covariance commonly used in spatial statistics \citep{Gneiting2012} is applied for the first time to model the turbulent background. 
Because of its flexibility, this covariance structure is shown to have more desirable features than previously-used stochastic models. 
Inertial oscillations are modeled with a complex-valued Ornstein-Uhlenbeck (OU) \citep{Arato1962, Jeffreys1968}, which provides a stochastic equivalent to an accepted set of coupled differential equations describing inertial oscillations. 
Further statistical innovation is employed in fitting the model with a variation on standard Whittle likelihood that compensates for both bias and aliasing and in allowing for non-stationarity through time-varying parameter estimates.
\par
The remainder of this report proceeds as follows.
 After a succinct review of spectral analysis, Section 2 discusses the model for the drifters and the methods used to fit it. 
 Section 3 presents the paper's results including analysis of a long time series with time-varying parameters and an application from a simulated model. 
We conclude the paper with a discussion of the value and limitations of the approach taken by \citet{Sykulski2016}.

@@

	\subsection{Spectral Analysis}
	\label{sec: specAnalysis}
	\indent We briefly review spectral analysis as all the analysis in \citet{Sykulski2016} is done in the spectral domain.   
	More thorough coverage of this material can be found in  \citep{Percival1993}, for example. 
	From a statistical perspective, a time series, $z(t)$, real or complex-valued,  can be understood as a realization of a stochastic process over time, where  $t = \{1,2,...,n\}$ index the observed time points.  
	Time series are often modeled as the sums of periodic functions, which, using Euler's formula, can be represented compactly as sums of appropriately-weighted complex exponentials. 
	Moreover, data in the time domain can easily be related to data in the frequency domain and vice versa using the Fourier or inverse Fourier transformations:
	\hdcomment{these are wrong} 	
	\begin{align}
z(t) = \int_{-\infty}^{\infty} f_{z}(\omega)e^{i\omega t}d\omega,&& f_{z}(\omega) = \int_{-\infty}^{\infty} z(t) e^{-i \omega t }dt.
\end{align}
\citep{Percival1993}. 
To understand the distribution of frequencies, the \textbf{\it{power spectral density}},
\begin{equation}
S(\omega) = \underset{T \rightarrow \infty}{\lim} \mathbb{E} \left(\frac{1}{2T} \left| \int_{-T}^{T} z(t) e^{-i \omega t}dt \right|^{2} \right),
\end{equation}
is often employed. 
This quantity represents the variance (or power) associated with each frequency \citep{Percival1993}. 
Power spectral densities are useful for statistical modeling and can be related to the autocovariance sequence of a time series using  Fourier transformation.  
We define the autocovariance  for $z(t)$ as $s(\tau) = \mathbb{E}[z(t) z^{*}(t - \tau)] $ where $z^{*}(t)$ represents the complex conjugate of z(t). Letting $\tau$ be the time lag, we have
\begin{align}
\label{eq: fourierPair}
S(\omega) = \int_{-\infty}^{\infty}s(\tau) e^{-i\omega t}d \tau \Longleftrightarrow s(\tau) = \frac{1}{2\pi} \int_{-\infty}^{\infty}S(\omega) e^{i \omega t} d\omega 
\end{align}

\noindent \citep{Sykulski2013}. For an observed time series, the power spectral density is often estimated with the periodogram, where 
\begin{equation}
\label{eq: perio}
\hat{S}(\omega) = \frac{\Delta}{N} \left( \sum_{t=1}^{N} z(t) e^{-i \omega t} \right)^{2}
\end{equation}
\citep{Sykulski2013}. 
Although intuitively appealing, this estimator has known problems that stem from  sampling a time series at discrete intervals. 
\textbf{\it{Aliasing}} refers to the fact that frequencies above the highest observable frequency cannot be captured, and, are instead, confounded with other frequencies in the spectrum.  
\hdcomment{Get this right!} Additionally, \textbf{\it{Leakage}} refers to the effects of because Fourier transforms are applied are a smooth, create discontinuities
$\textbf{\it{Tapering}}$ reduces leaking by \hdcomment{get this right!}
However, depending on how tapering is used, it can introduce its own  biases in estimation.

\subsection{Stochastic Model}
 To model a drifter's movement over time, the eastward and northward components of the drifter's velocity, denoted $u(t)$ and $v(t)$ respectively, are converted to a complex-valued velocity, $z(t) = u(t) + iv(t)$. 
 Separate stochastic models are introduced for the inertial oscillation and turbulent background components of the drifter's behavior. 
 Working in the frequency domain, the two components of the model are added together, creating a six-parameter spectral model. 
 Parameter estimates for this model are obtained via maximizing the model's blurred Whittle likelihood, a relatively new approximation to the true Whittle likelihood that compensates for bias and handles aliasing \citep{Sykulski2013}.

\subsubsection{Inertial oscillations}
We begin by discussing the model for the inertial oscillations. 
In a deterministic setting, inertial oscillations are often modeled with the following set of coupled differential equations
\begin{align}
\label{eq: deterOsc}
\frac{\partial u }{\partial t}  + f_{0} \nu &= F - cu,\\ \nonumber
\frac{\partial v}{\partial t} - f_{0}u &= G - cv,
\end{align}
where $u$ and $v$ again represent eastward and northward velocities, $f_{0}$ represents the inertial frequency in radians per unit time, and $F$ and $G$ are forces related to the wind  \citep{Pollard1970}. 
Suppressing the dependence on $t$ and using the complex representation of the velocity,  we show that these relationships can be equivalently expressed as follows
\begin{align}
\label{eq:diffEqDeriv}
\nonumber
\frac{dz}{dt} &= \frac{\partial u}{dt} + i\frac{\partial v(t)}{dt} \\ \nonumber
dz &= (F - c u- f_{0}v)dt + i(G - cv + f_{0}u)dt\\ \nonumber
&= f_{0}(-v + iu)dt - c(u + iv)dt + (F + iG)dt\\ \nonumber
&= if_{0}(u + iv)dt - c(u + iv)dt + (F + iG)dt\\ 
&= (if_{0} - c)z dt + (F + iG)dt. 
\end{align}
\citet{Sykulski2016} further replace the wind forcing term, $(F + iG)dt$  with complex-valued Brownian increments \citep{Mandelbrot1968} to obtain a stochastic analogue for Equation \ref{eq:diffEqDeriv},
 \begin{equation}
\label{eq: ouEq}
dz(t) = (i f_{0} -c) z(t) dt + A d Q(t). 
\end{equation}  
Equation \ref{eq: ouEq}  is known as the complex-valued Ornstein-Uhlenbeck (OU) stochastic differential equation. 
In general,  OU processes are considered continuous-time equivalents to the AR(1) process. Specifically,  the univariate OU process is the only mean-zero, stationary, Gaussian, and continuous Markov process, and, similarly the multivariate OU process is the only mean-zero $K$-variate process that is stationary, continuous, Gaussian, and Markov \citep{Schach1971}.
 The complex-valued OU process is the complex representation of a bivariate OU process.  
 In the formulation of Equation \ref{eq: ouEq}, $c > 0$ is a dampening parameter that forces the process to always eventually return to its mean position, $z = 0$.  
 The final term of Equation \ref{eq: ouEq} is a positive constant, $A$, multiplied by $dQ(t)$.  The term $Q(t), t \geq 0$ is a standard complex Wiener process,  meaning $Q(t) = Q_{1}(t) + i Q_{2}(t)$, where $Q_{1}(t)$ and $Q_{2}(t)$ are standard Wiener processes.
 We remind the reader that Wiener processes are the stochastic models originally used to represent Brownian motion. 
  Such a process $\{Q(t, \omega): t \geq 0\}$ on the measure space $(\omega, Q, P)$ must satisfy
\begin{enumerate}
\item $Q(0, \omega) = 0$ almost everywhere
\item $\{Q(t, \omega): t \geq 0\}$ is distributed normally on $(\Omega,Q, P)$
\item $Q(t + \tau, \omega) - Q(t, \omega)$ has mean 0 and variance $\tau$ for all $t, \tau > 0$
\end{enumerate}
\citep{Hida1980}.

To identify and estimate shifts in the angular frequency, presumably caused by eddies, \citet{Sykulski2016} propose replacing the known inertial requency, $f_{0}$,  in Equation \ref{eq: ouEq} with  a free parameter, $\omega_{0}$, to be estimated by the data. 
If the estimated angular frequency differs from the inertial frequency, the angular frequency of the eddy can be calculated as $\omega_{eddy} = f_{0} - \omega_{0}$, since $\omega_{0}$ is a sum of the inertial and eddy frequencies.

For time series modeling, we are interested in the autocovariance of this process at different time lags. To derive the autocovariance as given in Equation 4 of \citet{Sykulski2016}, we follow \citet{Arato1999} and consider the bivariate rather than complex-valued OU process.
This is possible since the autocovariance of a complex-valued random variable $z(t) = u(t) + i v(t)$ with time lag $\tau >0$ is defined as 
\begin{equation}
\label{eq: expec}
s(\tau) = \mathbb{E}[z(t) z^{*}(t + \tau)] = \mathbb{E}[u(t)u(t + \tau)] + \mathbb{E}[(v(t) v(t + \tau)] + i(\mathbb{E}[u(t) v(t + \tau)] -\mathbb{E}[u(t + \tau)v( t)])
\end{equation}
 \citep{DeIaco2003}. The differential equations for components $u(t)$ and $v(t)$ can be expressed in matrix form as 
\begin{align}
\left( \begin{array}{c} du(t) \\ dv(t) \end{array} \right) &= 
\left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right) \left( \begin{array}{cc} u(t) \\ v(t) \end{array} \right) + \left( \begin{array}{c} A d Q_{1}(t) \\ Ad Q_{2}(t) \end{array} \right). 
\end{align}
For a bivariate OU process, the covariance matrix at time lag $\tau > 0$ is 
\begin{align}
\label{eq: bivOU}
\boldsymbol{C}(\tau) = \left( \begin{array}{cc} \mathbb{E}[u(t + \tau)u(t)] & \mathbb{E}[u(t + \tau)v(t)] \\ 
\mathbb{E}[v(t + \tau)u(t)] & \mathbb{E}[v(t + \tau) v(t)] \end{array} \right) = \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\}\boldsymbol{C}(0)
\end{align}
where $\boldsymbol{C}(0)$ gives the covariance matrix at the initial time, $t$ \citep{Schach1971}.  \citet{Sykulski2016} specify that at the initial time, $z(t) \sim$ N(0, $A^{2}/2c$). We also note $u(t)$ and $v(t)$ are uncorrelated at the initial time, giving $\mathbb{E}[u(t)v(t)] = \mathbb{E}[v(t) u(t)] = 0$. Thus, Equation \ref{eq: expec} simplifies to
\begin{equation}
\label{eq: s0}
s(0) = \frac{A^{2}}{2c} =  \mathbb{E}[u(t)^{2}] + \mathbb{E}[(v(t)^{2}]. 
\end{equation}
Since $u(t)$ and $v(t)$ are assumed to be the same type of process, Equation \ref{eq: s0} implies $\mathbb{E}[u(t)^{2}] =\mathbb{E}[(v(t)^{2}] = A^{2}/4c$, which means Equation \ref{eq: bivOU} can be expressed as 
\begin{align}
\label{eq: covDeriv}
 \boldsymbol{C}(\tau) =  \exp \left\{ \left( \begin{array}{cc} -c & -\omega_{0} \\ \omega_{0} & -c \end{array} \right)\tau \right\} \left( \begin{array}{cc} A^{2}/4c & 0 \\ 0 & A^{2}/4c \end{array} \right)  &= 
\frac{A^{2}}{4c} \exp (-c \tau) \left( \begin{array}{cc} \cos \tau \omega_{0} & - \sin \tau \omega_{0} \\ \sin \tau \omega_{0} & \cos \tau \omega_{0}  \end{array} \right) 
\end{align}
where the second equality can be seen by expanding the infinite series representations for $\exp(-c \tau)$, $\sin(\tau \omega_{0})$, and $\cos( \tau \omega_{0})$. 
Finally, using the values obtained in Equation \ref{eq: covDeriv} in Equation \ref{eq: expec}, we obtain the autocovariance  as defined in \citet{Sykulski2016}
\begin{align}
\label{eq: ouAC} \nonumber
s^{(o)}(\tau) & = \frac{A^{2}}{4c} \exp(-c\tau) 2(\cos \tau \omega_{0} + \sin \tau \omega_{0}) \\
&= \frac{A^{2}}{2c} \exp(i \omega_{0}\tau) \exp(-c|\tau|).
\end{align}
 The $(o)$ superscript denotes that this equation is for the OU component of the full model. Using Equation \ref{eq: fourierPair}, the corresponding power spectral density can be found to be\begin{align}
\label{eq:ouPSD}
S^{(o)}(\omega) &=  \int_{-\infty}^{\infty} s^{(0)}(\tau) \exp (-i \omega \tau) d \tau = \frac{A^{2}}{(\omega - \omega_{0}) + c^{2}}. 
\end{align}

@@

Towards this end, the authors propose using a Mat\'{e}rn covariance structure \citep{Gneiting2012} for the turbulent background. 
The Mat\'{e}rn model has gained prominence in the spatial statistics literature due to its flexibility.
 In this context, the Mat\'{e}rn model proves useful, since it allows for both integer and fractional powers. 
 The Mat\'{e}rn model is defined either by its autocovariance 
\begin{align}
\label{eq:maternAC}
s^{(m)}(\tau) = \frac{B^{2}}{2^{\alpha - 1/2}\pi^{1/2} \Gamma(\alpha) h^{2 \alpha - 1}}(h|\tau|)^{\alpha - 1}\mathcal{K}_{\alpha - 1/2}(h|\tau|),
\end{align}
or power spectral density,
\begin{align}
\label{eq:maternPSD}
S^{(m)}(\omega) = \frac{B^{2}}{(\omega^{2} + h^{2})^{\alpha}},
\end{align}
where $\tau$ again denotes the time lag, $\Gamma(\alpha)$ is the gamma function,  $\mathcal{K}_{\eta}$ is a modified Bessel function of the second kind of order $\eta$ \citep{Stein2012}.  The superscript $(m)$ indicates that this is the Mat\'{e}rn part of the full model. 
The parameters to be estimated are the amplitude, $B$, the dampening, $h > 0$, and a smoothness parameter $\alpha > 1/2$.
 The parameter $\alpha$ is of particular importance, since it controls the smoothness, or degree of differentiability, of the process \citep{Fuentes2010}.

The authors emphasize that the turbulent background models previously proposed are special cases of the Mat\'{e}rn model. 
This means that these models can still be obtained under the current model if model fitting leads to their parameter estimate values. 
They also highlight that that Mat\'{e}rn model is more appropriate than the other obvious choice of a fractional power model, FBM, since FBM does not have a bound on the velocities and tends to drift quite far from its mean.
 Figure \ref{fig: fmbMat} illustrates these differences by comparing the paths of a simulated particle obtained with a Mat\'{e}rn  covariance versus a FBM covariance.

@@

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{ReplicatedFigures/fig2.pdf}
        \caption{Periodograms of the highlighted green and blue sections of the trajectories in Figure \ref{fig: fig1}. The inertial frequencies are marked in red. }
        	\label{fig: fig2}
\end{figure}
